\documentclass[10pt, a4paper]{article}
\input{../preamble.tex}


\begin{document}


%%% TD 1 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
{\tdtitle{1}{Set of real numbers}}

\begin{exercise}
  We say the set \(A\subset\R\) is \emph{bounded} provided \(A\) is
  both bounded above and bounded below.
  Prove that %\(A\) is bounded \emph{if and only if} \(\exists M>0\) such that \(\forall a\in A,~ \abs{a}\leq M\).
  \(
    \text{A is bounded}\iff
    \exists M>0,~\forall a\in A,~\abs{a}\leq M.
  \)
\end{exercise}

\begin{exercise}
  Determine whether the following sets are bounded below, or bounded above.
  (\textsf{Note:} You are allowed to use \(\floor{\cdot}\) in this exercise.)
  \begin{enumerate}
  \item \(A = \{\frac{n}{2}+1 \colon n\in \N\}\)
  \item \(B = \{\frac{n}{2}+1 \colon n\in \Z\}\)
  \item \(C = \{\frac{1}{n} \colon n\in\N\}\)
  \item \(D = \{ (-1)^n n \colon n\in\N\}\)
  \item \(E = \{ x\in\R \colon x^2\leq 3\}\)
    (\textsf{Indication:}
    For the last exercise even though it is tempting to use the
    \(\sqrt{\cdot}\) in order to solve for \(x\), you are advised \emph{not}
    to use it. This is because up until now we haven't showed that the
    \emph{number} \(\sqrt{3}\) exist or not.
    )
  \end{enumerate}
\end{exercise}

\begin{exercise}[Finding supremum and infimum]
  (\textsf{Note:} You are allowed to use \(\floor{\cdot}\) in this exercise.)
  \begin{enumerate}
  \item Let \(A = (-\infty, 5]\). Prove that \(\sup A = 5\).
  \item Let \(B = (2, \infty)\). Prove that \(\inf B = 2\).
  \item Let \(C = (1, 3]\). Prove that \(\sup C = 3\) and \(\inf C=1\).
  \item Let \(D = \{\frac{1}{n}\colon n\in\N\}\). Prove that \(\sup D=1\) and \(\inf D=0\).
  \end{enumerate}
\end{exercise}

\begin{exercise}
  Let \(x\in\R\) be a real number. Prove that there exists an integer
  \(n\in\N\) such that \(nx > 1\).
\end{exercise}
\begin{exercise}
  Let \(x,y\in\R\) be real numbers satisfying \(x<y\).
  Prove that there exists an integer \(n\in\N\) such that
  \(x + \frac1n < y.\)
\end{exercise}

\begin{exercise}[Existence of roots]
  \text{}
  \begin{enumerate}
  \item
    Let \(A = \{x\in\R \colon x^2 < 3\}\). Prove that \(A\) is bounded above and
    \((\sup A)^2 = 3\).
  \item 
    Let \(n\in\N\) and \(a\in\R\) with \(a > 0\).
    We denote the set  \(A = \{x\in\R \colon x^n < a\}\).
    Prove that \(A\) is bounded above and \((\sup A)^n = a\).
  \end{enumerate}
\end{exercise}

\begin{exercise}[Infimum]
  The goal of this exercise is to prove that any subset of
  \(\R\) that is bounded below has an infimum.
  Let \(A\subset\R\) be a subset that is bounded below.
  We denote
  \[B = \{-a~ \colon a\in A\}.\]
  \begin{enumerate}
  \item Prove that \(B\) is bounded above. Let \(\beta = \sup B\).
  \item Prove that \(\inf A = -\beta\). (Thus proving that infimum
    of any set that is bounded below exists.)
  \end{enumerate}
\end{exercise}

\begin{exercise}[Existence of integer part function]
  Let \(x\in\R\). Prove that there exists a unique integer \(N\in\Z\)
  satisfying \(N\leq x < N+1\). We denote this integer by \(N=\floor{x}\).
\end{exercise}

\begin{exercise}
  Let \(x,y\in\R\) with \(y-x>1\). Prove that there exists an integer
  \(m\in\Z\) such that \(x<m<y\).
\end{exercise}

\begin{exercise}
  Prove that the set \(A = \{\frac{p}{2^n}\colon p\in\Z,~n\in\N\}\)
  is dense in \(\R\).
\end{exercise}

\begin{exercise}
  Prove that the set of irrational numbers \(\R\smallsetminus\Q\)
  is dense in \(\R\).
\end{exercise}


%%% Solution
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \newpage
% \setcounter{exercise}{0}


% \begin{center}
%   {\large\sffamily Solution}
% \end{center}

% \begin{exercise}
%   We say the set \(A\subset\R\) is \emph{bounded} provided \(A\) is
%   both bounded above and bounded below.
%   Prove that 
%   \(
%     \text{A is bounded}\iff
%     \exists M>0,~\forall a\in A,~\abs{a}\leq M.
%   \)
% \end{exercise}
% \begin{proof}
%   todo.
% \end{proof}


% \begin{exercise}
%   Determine whether the following sets are bounded below, or bounded above.
%   \begin{enumerate}
%   \item \(A = \{\frac{n}{2}+1 \colon n\in \N\}\)
%   \item \(B = \{\frac{n}{2}+1 \colon n\in \Z\}\)
%   \item \(C = \{\frac{1}{n} \colon n\in\N\}\)
%   \item \(D = \{ (-1)^n n \colon n\in\N\}\)
%   \item \(E = \{ x\in\R \colon x^2\leq 3\}\)
%   \end{enumerate}
% \end{exercise}
% \begin{proof}
%   todo
% \end{proof}




\end{document}
